Feature importance analysis

### 2020-03-04 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE, verbose = TRUE)

  df_wavelengths_from_indices
)
```

@@

data_hr_merged <- left_join(spectra_df, fi_ranked_hr, by = c("wavelength")) %>%
  mutate(class = "HR") %>%
  mutate(reflectance = as.numeric(.data$reflectance))
```

@@

```{r feature-importance-5, echo=FALSE}
plot_reflectance_imp <- function(data) {
  ggplot(data, aes(x = .data[["wavelength"]], y = .data[["importance"]])) +
    geom_line(aes(
      x = .data[["wavelength"]], y = reflectance,
      color = "Spectral Signature of Vegetation"
    ),
    linetype = "solid",
    size = 0.6
    ) +
    # HR: plot only rank 11:last
    geom_point(
      data = data_hr_merged[which(!(data_hr_merged$rank %in% c(1:10))), ],
      aes(color = "HR: 11:122"),
      size = 1.8,
      shape = 5
    ) +
    # HR: plot only rank 1:10
    geom_point(
      data = data_hr_merged[which(data_hr_merged$rank %in% c(1:10)), ],
      aes(color = "HR: 1:10"),
      size = 1.8,
    ) +

    # VI: rank 1:10
    geom_point(
      data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
      aes(x = wavelength, y = importance, group = class),
      size = 2, shape = 4, color = "darkgreen"
    ) +
    # VI: rank 1:10
    geom_line(
      data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
      aes(x = wavelength, y = importance, group = class, color = "VI: 1:10"),
      size = 0.2, linetype = "dashed"
    ) +

    # VI: rank 11:89
    geom_point(
      data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
      aes(x = wavelength, y = importance, group = class),
      color = "grey",
      shape = 5, size = 2
    ) +
    # # VI: rank 11:89
    geom_line(
      data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
      aes(x = wavelength, y = importance, group = class, color = "VI: 11:89"),
      size = 0.2, linetype = "dashed"
    ) +

    scale_color_manual(values = c(
      "HR: 11:122" = "grey",
      "HR: 1:10" = "black",
      "Spectral Signature of Vegetation" = "black",
      "VI: 1:10" = "darkgreen",
      "VI: 11:89" = "grey"
    )) +
    scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
    scale_y_continuous(
      sec.axis = sec_axis(~ scale(-.,
        center = FALSE,
        scale = max(., na.rm = TRUE)
      ),
      labels = c(1.0, 0.75, 0.55, 0.25, 0),
      name = "Scaled Reflectance [0, 1]"
      )
    ) +
    guides(color = guide_legend(
      title = NULL,
      override.aes = list(
        linetype = c("blank", "blank", "solid", "dashed", "dashed"),
        shape = c(19, 5, NA, 4, 4),
        color = c("black", "grey", "black", "darkgreen", "grey")
      )
    )) +
    labs(
      title = "Spatial Variable Importance",
      subtitle = paste0(
        "Top ten features of 'VI' and 'HR' datasets along a simulated spectral signature of vegetation.\n",
        "Labels show the hyperspectral band number for dataset 'HR' and the index name for dataset 'VI'."
      ),
      caption = "Learner: SVM; 100 Monte-Carlo Iterations",
      y = "Mean decrease in RMSE", x = "Wavelength [nm]"
    ) +
    ggrepel::geom_label_repel(
      # only label the best 10 features
      data = df_label_all,
      label = df_label_all$class,
      nudge_x = 0.1, nudge_y = 0.1,
      size = 3
    ) +
    ggpubr::theme_pubclean()
}

@@

```{r fi-permut-vi-hr, dev = c("png", "pdf"), warning=FALSE}
p1 <- data_hr_merged %>%
  plot_reflectance_imp()
p1
```

>>>>

  df_wavelengths_from_indices,
  spec_sigs
)
```

@@

data_hr_merged <- left_join(spectra_df, fi_ranked_hr, by = c("wavelength")) %>%
  left_join(spec_sigs, by = c("wavelength")) %>% 
  mutate(class = "HR") %>%
  mutate(reflectance = as.numeric(reflectance))
```

@@

```{r feature-importance-5, echo=FALSE}
plot_reflectance_imp_hr <- function(data) {
  ggplot(data, aes(x = .data[["wavelength"]], y = .data[["importance"]])) +
    geom_line(
      data = data_hr_merged[!is.na(data_hr_merged$laukiz1), ],
      aes(
      x = wavelength, y = laukiz1,
      color = "Spectral Signature of Laukiz1"
    ),
    na.rm = TRUE,
    linetype = "solid",
    size = 0.6
    ) +
    geom_line(
      data = data_hr_merged[!is.na(data_hr_merged$laukiz2), ],
      aes(
      x = wavelength, y = laukiz2,
      color = "Spectral Signature of Laukiz2"
    ),
    na.rm = TRUE,
    linetype = "solid",
    size = 0.6
    ) +
    geom_line(
      data = data_hr_merged[!is.na(data_hr_merged$luiando), ],
      aes(
      x = wavelength, y = luiando,
      color = "Spectral Signature of Luiando"
    ),
    na.rm = TRUE,
    linetype = "solid",
    size = 0.6
    ) +
    geom_line(
      data = data_hr_merged[!is.na(data_hr_merged$oiartzun), ],
      aes(
      x = wavelength, y = oiartzun,
      color = "Spectral Signature of Oiartzun"
    ),
    na.rm = TRUE,
    linetype = "solid",
    size = 0.6
    ) +

    # HR: plot only rank 11:last
    geom_point(
      data = data_hr_merged[which(!(data_hr_merged$rank %in% c(1:10))), ],
      aes(color = "HR: 11:122"),
      size = 1.8,
      shape = 1,
      na.rm = TRUE
    ) +
    # HR: plot only rank 1:10
    geom_point(
      data = data_hr_merged[which(data_hr_merged$rank %in% c(1:10)), ],
      aes(color = "HR: 1:10"), 
      shape = 20,
      size = 1.8,
      na.rm = TRUE
    ) +

    scale_color_manual(values = c(
      "HR: 11:122" = "grey",
      "HR: 1:10" = "black",
      "Spectral Signature of Laukiz1" = "#BC3C29",
      "Spectral Signature of Laukiz2" = "#0072B5",
      "Spectral Signature of Luiando" = "#E18727",
      "Spectral Signature of Oiartzun" = "#20854E"
    )) +
    scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
    scale_y_continuous(
      sec.axis = sec_axis(~ scale(-.,
        center = FALSE,
        scale = max(., na.rm = TRUE)
      ),
      labels = c(1.0, 0.75, 0.55, 0.25, 0),
      name = "Scaled Reflectance [0, 1]"
      )
    ) +
    guides(color = guide_legend(
      title = NULL,
      override.aes = list(
        linetype = c("blank", "blank", "solid", "solid", "solid" ,"solid"),
        shape = c(20, 1, NA, NA, NA, NA),
        color = c("black", "grey", "#BC3C29", "#0072B5", "#E18727", "#20854E")
      )
    )) +
    labs(
      title = "Permutation-based Variable Importance for dataset 'HR'",
      subtitle = paste0(
        "The ten most important features are labeled by their band number."
      ),
      caption = "Learner: SVM; 100 Monte-Carlo Iterations",
      y = "Mean decrease in RMSE", x = "Wavelength [nm]"
    ) +
    ggrepel::geom_label_repel(
      # only label the best 10 features
      data = df_wavelengths_from_hr,
      label = df_wavelengths_from_hr$class,
      nudge_x = 0.1, nudge_y = 0.1,
      size = 3,
      na.rm = TRUE
    ) +
    ggpubr::theme_pubclean()
}

plot_reflectance_imp_vi <- function(data) {
  ggplot(data, aes(x = .data[["wavelength"]], y = .data[["importance"]])) +
    geom_line(
      data = data_hr_merged[!is.na(data_hr_merged$laukiz1), ],
      aes(
      x = wavelength, y = laukiz1,
      color = "Spectral Signature of Laukiz1"
    ),
    na.rm = TRUE,
    linetype = "solid",
    size = 0.6
    ) +
    geom_line(
      data = data_hr_merged[!is.na(data_hr_merged$laukiz2), ],
      aes(
      x = wavelength, y = laukiz2,
      color = "Spectral Signature of Laukiz2"
    ),
    na.rm = TRUE,
    linetype = "solid",
    size = 0.6
    ) +
    geom_line(
      data = data_hr_merged[!is.na(data_hr_merged$luiando), ],
      aes(
      x = wavelength, y = luiando,
      color = "Spectral Signature of Luiando"
    ),
    na.rm = TRUE,
    linetype = "solid",
    size = 0.6
    ) +
    geom_line(
      data = data_hr_merged[!is.na(data_hr_merged$oiartzun), ],
      aes(
      x = wavelength, y = oiartzun,
      color = "Spectral Signature of Oiartzun"
    ),
    na.rm = TRUE,
    linetype = "solid",
    size = 0.6
    ) +

    # VI: rank 1:10
    geom_point(
      data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
      aes(x = wavelength, y = importance, group = class),
      size = 1.4, shape = 20
    ) +
    # VI: rank 1:10
    geom_line(
      data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
      aes(x = wavelength, y = importance, group = class, color = "VI: 1:10"),
      size = 0.2, linetype = "dashed"
    ) +

    # VI: rank 11:89
    geom_point(
      data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
      aes(x = wavelength, y = importance, group = class),
      color = "grey",
      shape = 1, size = 2
    ) +
    # VI: rank 11:89
    geom_line(
      data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
      aes(x = wavelength, y = importance, group = class, color = "VI: 11:89"),
      size = 0.2, linetype = "dashed"
    ) +

    scale_color_manual(values = c(
      "VI: 11:89" = "grey",
      "VI: 1:10" = "black",
      "Spectral Signature of Laukiz1" = "#BC3C29",
      "Spectral Signature of Laukiz2" = "#0072B5",
      "Spectral Signature of Luiando" = "#E18727",
      "Spectral Signature of Oiartzun" = "#20854E"
    )) +
    scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
    scale_y_continuous(
      sec.axis = sec_axis(~ scale(-.,
        center = FALSE,
        scale = max(., na.rm = TRUE)
      ),
      labels = c(1.0, 0.75, 0.55, 0.25, 0),
      name = "Scaled Reflectance [0, 1]"
      )
    ) +
    guides(color = guide_legend(
      title = NULL,
      override.aes = list(
        linetype = c("solid", "solid", "solid" ,"solid","dashed", "dashed"),
        shape = c(NA, NA, NA, NA, 20, 1),
        color = c("#BC3C29", "#0072B5", "#E18727", "#20854E", "black", "grey")
      )
    )) +
    labs(
      title = "Permutation-based Variable Importance for dataset 'VI'",
      subtitle = paste0(
        "The ten most important features are labeled by their index name."
      ),
      caption = "Learner: SVM; 100 Monte-Carlo Iterations",
      y = "Mean decrease in RMSE", x = "Wavelength [nm]"
    ) +
    ggrepel::geom_label_repel(
      # only label the best 10 features
      data = df_wavelengths_from_indices_label,
      label = df_wavelengths_from_indices_label$class,
      nudge_x = 0.1, nudge_y = 0.1,
      size = 3,
      na.rm = TRUE
    ) +
    ggpubr::theme_pubclean()
}

@@

```{r fi-permut-hr, warning=FALSE, fig.show='hide'}
p11 <- data_hr_merged %>%
  plot_reflectance_imp_hr()
p11
```

```{r fi-permut-vi, warning=FALSE, fig.show='hide'}
p12 <- data_hr_merged %>%
  plot_reflectance_imp_vi()
p12
```

```{r fi-permut-vi-hr, dev = c("png", "pdf"), warning=FALSE, fig.asp=1.1}
p11 / p12
```
